Turbo fast no compromise gravitational wave parameter estimation

### 2023-01-20 src/tex/ms.tex
It should compiles now
edits up to and inc Sec. 2.5

The main objective in PE is to obtain a multidimensional posterior distribution
$p(\mathbf{\theta} \mid d)$ on parameters $\mathbf{\theta}$ given strain
data $d$.  Such probability density represents our best inference of
the source properties, an encodes all relevant information contained in the
observed data.
To compute this object, we use Bayes' theorem to write
\begin{align}
    p(\theta \mid d) = \frac{\mathcal{L}(d \mid \theta)\pi(\theta)}{p(d)}\, ,
\end{align}
where $\mathcal{L}(d \mid \theta)$ is the likelihood function, $\pi(\theta)$ is the
prior distribution, and $p(d)$ is the evidence. Since the evidence is a
normalization constant that does not depend on the source parameters, it is
often omitted if we are only interested in the posterior distribution. The prior
distribution is often chosen to be some simple distribution, such as uniformly
distributed in the component masses or a Gaussian distribution in the spins, or it could encode astrophysical information.
Assuming the noise is drawn from a Gaussian process, the
log-likelihood for GW data is given by
\begin{align}
    \log{\mathcal{L}(d \mid \theta)} = -\frac{1}{2} \left\langle d-h(\mathbf{\theta}) \mid d-h(\mathbf{\theta})\right\rangle,
\label{eq: loglikelihood}
\end{align}
where $d$ is the observed strain data, $h(\mathbf{\theta})$ is the signal predicted by a
waveform model with a specific set of source parameters $\theta$. The right hand side of
Eq.~\eqref{eq: loglikelihood} can be evaluated in either the time or frequency domains.
For stationary noise, it is computationally cheaper to compute the likelihood in the frequency
domain, and the noise-weighted inner product can be written 
\begin{align}
    % \left<a|b\right> = 4 Re\int \frac{a^*(f)b(f)}{\mathcal{S}_n(f)} df,
    \left\langle a \mid b\right\rangle = 4 \Re \int \frac{a^*(f)b(f)}{\mathcal{S}_n(f)}\, \mathrm{d}f \, ,
\label{eq: innerproduct}
\end{align}
where $\mathcal{S}_n(f)$ is the noise power spectral density (PSD).
In practice, the integral becomes a discrete sum over a finite number of
samples determined by the sampling rate of the detector data and duration of
the observation.

@@

Given eq. \ref{eq: loglikelihood} and the prior, one can evaluate the posterior
density function over the entire parameter space of interest to obtain the most
probable set of points that are consistent with the data. However, direct
sampling quickly becomes intractable as the dimensionality of the parameter
space increases beyond a few dimensions. Markov chain Monte Carlo (MCMC)
\cite{gelmanbda04} is a common method employed to generate samples when direct
sampling is not possible.

In MCMC, the posterior distribution is approximated by a Markov chain that
converges to the target distribution \cite{10.1214/aos/1176325750}. The Markov
chain is constructed by iteratively proposing a new point in the parameter space
based on the current location of the chain. The proposed point is accepted with
a probability that is usually set to be proportional to the ratio of the
posterior density evaluated at the proposed point and the current point. The
chain can either accept the proposal and move to the new location or reject the
proposal and stay at the current location. This process is repeated until the
chain converges to the target distribution. The samples generated by the chain
are then used to estimate the quantities of interest, such as the mean and
credible intervals of the source parameters. In practice, since we do not know
the target distribution ahead of time, the MCMC process is usually repeated
until a certain criterion is met, such as a Gelman-Rubin convergence statistic
\cite{10.2307/2246093} lower than a threshold or simply after a certain number
of iterations.

Compared to direct sampling, MCMC algorithms only explore regions that are
highly probable, thus reducing the computational cost by not wasting resources in
regions where it is unlikely to generate the data. However, MCMC algorithms come
with their own set of issues. To illustrate what difficulties an MCMC algorithm
may face, we can examine one of the most vanilla MCMC algorithm, the
Metropolis-Hastings algorithm with a Gaussian kernel. Starting at some initial
point, one can draw a proposed point from a Gaussian transition kernel, defined
as
\begin{align}
    q(\mathbf{x},\mathbf{x_0})= \mathcal{N}(\mathbf{x}|\mathbf{x_0},\mathbf{C}),
\end{align}
where $\mathbf{x_0}$ is the current location of the chain, $\mathbf{x}$ is the
proposed location, and $\mathbf{C}$ is the covariance matrix of the Gaussian. In
the simplest case, we can pick $\mathbf{C}$ to be a diagonal matrix with a
constant value, which corresponds to an isotropic Gaussian center around the
current location with a constant variance. And the acceptance criteria is
defined as
\begin{align}
\alpha(\mathbf{x},\mathbf{x_0}) = \min\left(1,\frac{p(\mathbf{x})q(\mathbf{x_0},\mathbf{x})}{p(\mathbf{x_0})q(\mathbf{x},\mathbf{x_0})}\right).
\label{eq:Gaussian_acceptance}
\end{align}

We can see from eq. \ref{eq:Gaussian_acceptance} the acceptance rate is
proportion to the fraction of volume where the posterior density at the proposed
location is higher than the current location within the Gaussian transition
kernel. If we choose the variance of the transition kernel to be too large, this
fraction will be small hence the acceptance rate will be poor. On the other
hand, if one chooses the variance to be too small, nearby samples will be
correlated. In both cases, the efficiency in constructing the chain with a
target number of independent samples is suboptimal. There is often a tuning
process before we run the MCMC algorithm to find the optimal tuning parameters,
in this example the variance of the Gaussian, to ensure close to optimal
performance.

However, as we are dealing with higher dimensional problems, even the optimally
tuned Gaussian transition kernel does not guarantee good performance. In order
to have a reasonable acceptance rate, the variance of the Gaussian has to be
smaller in a higher dimensional space, which means the transition kernel is in
general making smaller and smaller steps as we increase the dimensionality
\cite{2017arXiv170102434B}.

Transition kernels that leverage gradient information of the target
distribution can help to address this issue of shortening steps in a high
dimensional space. Instead of proposing a new point by drawing from a Gaussian,
one can use the gradient evaluation at the current location to propose a new
point. For example, Metropolis-adjusted Langevin algorithm (MALA) \cite{10.2307/2346184}
place a unit Gaussian at the tip of the gradient vector at the current position,
\begin{align}
    \mathbf{x} = \mathbf{x_0} + \tau \nabla\log{p(\mathbf{x_0})} + \sqrt{2\tau}N(0,\mathbf{I}),
\end{align}
where $\tau$ is the step size, which is a tuning parameter. Compared to a
Gaussian centers at the current location, the MALA transition kernel is more
likely to propose a point into the higher posterior density region because of
the gradient term, which helps to boost the acceptance rate.

While transition kernels that use gradient information can help to improve the
acceptance rate, computing the gradient of the posterior density function
introduce additional computational cost, which is not necessarily beneficial in
terms of sampling time. If one wants to compute the gradient information through
finite differencing, the additional computation cost goes as at least $\sim
\mathcal{O}(2n)$, where $n$ is the dimension of the problem. \texttt{Jax} allows
us to compute the gradient of the likelihood function with respect to the
parameters through automatic differentiation, which gives the gradient
information down to machine precision at around the same order of time compared
to evaluating the posterior itself. Having access to gradient information
through automatic differentiation is crucial to make the trade-off between using
gradient-based transition kernels and additional computation cost favorable.

@@

% Problem with just HMC or MALA
While gradient-based samplers have been shown to be superior in terms of
performance when compared to other gradient-free algorithms in many practical
examples, there are still a number of classes of problems gradient-based
samplers do not solve well. \footnote{To be specific, we are referring to
sampling algorithms that use first-order derivatives here. Sampling algorithms
that use the information of higher order derivatives such as manifold-MALA and
Riemannian-HMC \cite{RMHMC} can in principle decorrelate local correlation in
the target distribution, however, they often have instability issues when they
are used on real-life applications, so they are not used often in practice.} For
example, target distribution that exhibits local correlation is hard to deal
with, since by construction first order gradient-based algorithms can only
handle global correlation structure \cite{2017arXiv170102434B}. Another example
is multi-modality. If there are multiple modes in the target distribution, an
independent chain will likely be trapped in one mode and take an extremely long
time to transverse between the modes \cite{2018arXiv180803230M}. This means the
relative weights between modes will much longer to sample compared to the shape
of each mode.

% Long burn-in too
Moreover, before we can use the sampling chain to estimate the quantity we care about,
the sampler often needs to first find the most probable region in the target
space, which is a common process that is referred to as "burn-in" in the
literature. This means one would discard a certain amount of data generated from
the beginning and only use the later part to estimate the quantities of
interest. The burn-in phase of a gradient-based sampler is often as long as the
sampling phase, which means a good portion of computation is not necessarily
helpful in estimating the target quantities.

% Crux of normalizing flow
Normalizing flow is a neural-network based technique that aims at learning a
mapping between a simple distribution, such as a Gaussian distribution, to a
complex distribution, often given in the form of data samples from a target
distribution \cite{2019arXiv190809257K, 2019arXiv191202762P}. Once the network is trained, one can evaluate the probability
density of the complex distribution and samples from it very efficiently.
The core equation of normalizing flow is given by
\begin{align}
    p_x(X) = p_z(Z) \left| \frac{\partial f}{\partial z}\right|,
\end{align}
where $p_x(X)$ is the complex target distribution $p_z(Z)$ is the simple latent
distribution and $f$ is an invertible parameterized transform that
connect the two distributions. For a detailed discussion of the algorithm, we
refer the readers to \cite{2019arXiv190809257K, 2019arXiv191202762P}.

% The flowMC algorithm
As mentioned in the previous subsection, one of the main issue gradient-based
sampler is it does not explore locally correlated features and multi-modality well. This
is exactly where the normalizing flow model can help. Once multiple independent
chains have generated samples, we can combine the samples and feed them to the
normalizing flow, which learns the global landscape of the target
distribution. Note that since the independent chains are generated with the same
number of samples, the normalization between posterior density represented by
different chains is in general different from the true target distribution. To
adjust for this inaccuracy, we use the normalizing flow model as a
\textit{proposal} distribution. Given a trained normalizing flow model, we can
generate the proposed jump in the target space by sampling from the latent
distribution $z \sim p_z(Z)$, then push it through the learned map given by the
normalizing flow model $x=f(z)$. The acceptance criterion is given by
\begin{align}
    \alpha(\mathbf{x},\mathbf{x_0}) = \min \left[ 1, \frac{\hat{\rho}(\mathbf{x_0})\rho_*(\mathbf{x})}{\hat{\rho}(\mathbf{x})\rho_*(\mathbf{x_0})}\right],
\end{align}
where $\hat{\rho}$ is the probability density estimated by the normalizing flow
model, $\rho_*$ is the probability density evaluated using the target function,
and $x_0$ is the current position. We can see the flow distribution is the
target distribution when the accepting probability is 1. When the normalizing
flow model has not converged to the target distribution, only a portion of the
proposed jumps will be accepted. This means an MCMC process using the
normalizing flow model as the proposal distribution can adjust the normalization
across different regions of the target parameter space by rejecting jumps into
less likely regions. The training and sampling are then repeated until certain
criteria are met. Note that every time we retrain the network, we are breaking
the Markov properties since we are changing the proposal distribution. To
produce final samples that can be used to estimate target quantities, one has to
freeze the normalizing flow model and not retrain during the final sampling
phase in order to satisfy the detailed balance condition. We use the package
\texttt{flowMC} \cite{2022arXiv221106397W}. The pseudocode of the algorithm is
given in Algorithm \ref{alg:cap}.

\begin{algorithm}
\caption{flowMC pseudocode}\label{alg:cap}
\KwIn{Initial position $ip$}
\Parameters{n training loop $nt$, n production loop $np$}
\Variables{current chain $cc$, current position $cp$, current NF parameters $\Theta$}
\KwResult{Chains}
$cp \leftarrow ip$\\
\tcc{Training loop}
    \For{$i<nt$}{
        $cc, cp \leftarrow LocalSampling(cp)$\\
        $\Theta \leftarrow TuneNF(cc)$\\
        $chains, cp \leftarrow GlobalSampling(cp, \Theta)$ \\
        $cc \leftarrow Append(cc, chains)$
    }
\tcc{Production loop}
    \For{$i<np$}{
        $c_{local}, cp \leftarrow LocalSampling(cp)$\\
        $c_{global}, cp \leftarrow GlobalSampling(cp, \Theta)$ \\
        $Chains \leftarrow Append(Chains, c_{local}, c_{global})$
    }

@@

Modern accelerators such as GPUs and TPUs are designed to execute large-scale
dense computation. They are often much more cost-efficient than using many CPUs
when it comes to solving problems that can be benefited from parallelization.
The downside of these accelerators compared to CPUs is they can only perform a
more restricted set of operations and are often less performant when they are
dealing with serial problems. Parameter estimation with MCMC is a serial
problem since each new sample generated from a chain depends on the last sample
in the chain. This means naively putting the problem on an accelerator is more
likely to harm the performance instead of improving it.

% More sample helps training normalizing flow
In this work, the use of accelerators provides two independent perks that
tremendously benefit the parameter estimation process. First, using
accelerators allow us to run many independent chains at the same time, which
benefits the training of the normalizing flow. Since we generate the data we use
to train the normalizing flow on the fly, the more independent data we can feed
to the training process, the higher chance the normalizing flow can learn a
reasonable representation of the global landscape of the target distribution. If
we only use a small number of chains, we are limited to the correlated samples
from each chain, so we have to run more sequential steps to get the same amount
of independent data as compared to running more chains, where the former option does
not benefit from parallelization but the latter does. In another word, being
able to use many independent chains help the normalizing flow learn the global
landscape faster in Wall time.

% GPU helps packing a shit ton of waveform evaluation
Another benefit accelerator brings is the parallel evaluation of waveforms. Since
the waveform models being used in PE are approximants, they can be evaluated at
any given time or frequency independent of what has been generated before the
time or frequency. This means predicting the waveform can be trivially
parallelized over frequency. Together with heterodyned likelihood, we can
evaluate the likelihood at $\mathcal{O}(10^7)$ different locations on an Nvidia
A100 GPU. The high throughput of likelihood evaluations unlocks the potential of
the \texttt{flowMC} sampling algorithms.

>>>>

The main objective in PE is to obtain a multidimensional posterior distribution
$p(\mathbf{\theta} \mid d)$ on parameters $\mathbf{\theta}$ given strain
data $d$.  Such probability density represents our best inference of
the source properties, an encodes all relevant information contained in the
observed data.
To compute this object, we use Bayes' theorem to write
\begin{align} \label{eq: bayes}
    p(\theta \mid d) = \frac{\mathcal{L}(d \mid \theta)\pi(\theta)}{p(d)}\, ,
\end{align}
where $\mathcal{L}(d \mid \theta)$ is the likelihood function, $\pi(\theta)$ is the
prior distribution, and $p(d)$ is the evidence. Since the evidence is a
normalization constant that does not depend on the source parameters, it is
often omitted if we are only interested in the posterior distribution. The prior
distribution is often chosen to be some simple distribution, such as uniformly
distributed in the component masses or a Gaussian distribution in the spins, or it could encode astrophysical information.
Assuming the noise is drawn from a Gaussian process, the
log-likelihood for GW data is given by
\begin{align}
    \log{\mathcal{L}(d \mid \theta)} = -\frac{1}{2} \left\langle d-h(\mathbf{\theta}) \mid d-h(\mathbf{\theta})\right\rangle,
\label{eq: loglikelihood}
\end{align}
where $d$ is the observed strain data, $h(\mathbf{\theta})$ is the signal predicted by a
waveform model with a specific set of source parameters $\theta$. The right hand side of
Eq.~\eqref{eq: loglikelihood} can be evaluated in either the time or frequency domains.
For stationary noise, it is computationally cheaper to compute the likelihood in the frequency
domain, and the noise-weighted inner product can be written 
\begin{align}
    % \left<a|b\right> = 4 Re\int \frac{a^*(f)b(f)}{\mathcal{S}_n(f)} df,
    \left\langle a \mid b\right\rangle = 4 \Re \int \frac{a^*(f)b(f)}{\mathcal{S}_n(f)}\, \mathrm{d}f \, ,
\label{eq: innerproduct}
\end{align}
where $\mathcal{S}_n(f)$ is the noise power spectral density (PSD).
In practice, the integral becomes a discrete sum over a finite number of
samples determined by the sampling rate of the detector data and duration of
the observation.

@@

Given Eq.~\eqref{eq: loglikelihood} and the prior, one can evaluate the
posterior density function, Eq.~\eqref{eq: bayes}, over the entire parameter
space of interest to obtain the most probable set of values that are consistent
with the data. However, directly sampling this posterior quickly becomes
intractable as the dimensionality of the parameter space increases beyond a few
dimensions. Markov chain Monte Carlo \cite{gelmanbda04} is a common
method employed to generate samples from the target posterior when direct
sampling is not possible.

In MCMC, the posterior distribution is approximated by a Markov chain that eventually
converges to the target distribution \cite{10.1214/aos/1176325750}. The Markov
chain is constructed by iteratively proposing a new point in the parameter space
based on the current location of the chain. The proposed point is accepted with
a probability that is usually set to be proportional to the ratio of the
posterior density evaluated at the proposed point and the current point. The
chain can either accept the proposal and move to the new location, or reject the
proposal and stay at the current location. This process is repeated until the
chain converges to the target distribution. The samples generated by the chain
are then used as a fair sample to estimate the quantities of interest, such as the mean and
credible intervals of the source parameters. In practice, since we do not know
the target distribution ahead of time, the MCMC process is usually repeated
until a certain criterion is met, such as a Gelman-Rubin convergence statistic
\cite{10.2307/2246093} lower than certain threshold, or simply after a fixed number
of iterations.

Compared to direct sampling, MCMC algorithms only explore regions that are
highly probable, thus reducing the computational cost by not wasting resources
in regions where it is unlikely to generate the observed data. However, MCMC
algorithms come with their own set of issues. To illustrate what difficulties
MCMC may face, we can examine one of the most vanilla MCMC algorithms: the
Metropolis-Hastings algorithm with a Gaussian kernel. Starting at some initial
point, one can draw a proposed point from a Gaussian transition kernel, defined
as
\begin{align}
    q(\mathbf{x},\mathbf{x_0})= \mathcal{N}(\mathbf{x} \mid \mathbf{x_0},\mathbf{C}),
\end{align}
where $\mathbf{x_0}$ is the current location of the chain, $\mathbf{x}$ is the
proposed location, and $\mathbf{C}$ is the covariance matrix of the Gaussian. In
the simplest case, we can pick $\mathbf{C}$ to be a diagonal matrix with a
constant value, which corresponds to an isotropic Gaussian center around the
current location and with a fixed variance. The acceptance criterion is
defined as
\begin{align}
\alpha(\mathbf{x},\mathbf{x_0}) = \min\left(1,\frac{p(\mathbf{x})q(\mathbf{x_0},\mathbf{x})}{p(\mathbf{x_0})q(\mathbf{x},\mathbf{x_0})}\right).
\label{eq:Gaussian_acceptance}
\end{align}

We can see from Eq.~\eqref{eq:Gaussian_acceptance} that the acceptance rate is
proportional to the fraction of volume where the posterior density at the
proposed location is higher than the current location within the Gaussian
transition kernel. If we choose the variance of the transition kernel to be too
large, this fraction will be small hence the acceptance rate will be poor. On
the other hand, if one chooses the variance to be too small, nearby samples
will be correlated, and it will take a long time for the chain to wander. In
both cases, the efficiency in constructing the chain with a target number of
independent samples is suboptimal. Consequently, there is often a tuning
process before we run the MCMC algorithm to find the optimal settings for the
algorithm (in this example, the variance of the Gaussian) to ensure the best
possible performance.

However, as we often deal with high-dimensional problems, even the optimally
tuned Gaussian transition kernel does not guarantee good performance. In order
to have a reasonable acceptance rate, the variance of the Gaussian has to be
smaller in a higher dimensional space, which means that the transition kernel will generally make smaller and smaller steps as we increase the dimensionality of the problem
\cite{2017arXiv170102434B}.

Transition kernels that leverage gradient information of the target
distribution can help to address this issue of shortening steps in a high
dimensional space. Instead of proposing a new point by drawing from a Gaussian,
one can use the gradient evaluated at the current location to propose a new
point, so that the evolution of the chain is preferentially directed to regions
of higher probability. For example, Metropolis-adjusted Langevin algorithm
(MALA) \cite{10.2307/2346184} place a unit Gaussian at the tip of the gradient
vector at the current position,
\begin{align}
    \mathbf{x} = \mathbf{x_0} + \tau \nabla\log{p(\mathbf{x_0})} + \sqrt{2\tau}N(0,\mathbf{I}),
\end{align}
where $\tau$ is a step size chosen during the tuning stage. Compared to a
Gaussian centered at the current location, the MALA transition kernel is more
likely to propose a point in the higher posterior density region because of
the gradient term, which helps boost the acceptance rate.

While transition kernels that use gradient information can help improve the
acceptance rate, computing the gradient of the posterior density function
introduces an additional computational cost, which is not necessarily
beneficial in terms of sampling time. If one wants to compute the gradient
through finite differencing, the additional computational cost goes as at least
${\sim} \mathcal{O}(2n)$, where $n$ is the dimension of the problem. On the
other hand, automatic differentiation schemes like \textsc{Jax} allow us to
compute the gradient of the likelihood function with respect to the parameters
through automatic differentiation, which gives the gradient information down to
machine precision at around the same order of time compared to evaluating the
posterior itself. Thus, having access to gradient information through automatic
differentiation is crucial to making gradient-based transition kernels
favorable in terms of computing cost.

@@

% Problem with just HMC or MALA
While gradient-based samplers have been shown to outperform gradient-free
algorithms in many practical applications, there remain classes of problems that
most gradient-based samplers do not solve well. For example, first-order
gradient-based algorithms struggle with target distributions that exhibit local
correlations, since they can only handle global structures by construction
\cite{2017arXiv170102434B}.\footnote{Sampling algorithms that use the
information of higher order derivatives such as manifold-MALA and
Riemannian-HMC \cite{RMHMC} can in principle handle local correlations in the
target distribution; however, they often encounter instabilities when used in
real-life applications, so their use is rare practice.} Another example is
multi-modality: if there are multiple modes in the target distribution,
individual chains will likely be trapped in one mode and take an extremely long
time to transverse between the modes \cite{2018arXiv180803230M}.  This means
that the relative weights between modes will take much longer to estimate than
the shape of each mode.

% Long burn-in too
Moreover, before we can use the sampling chain to estimate the posterior
quantities we care about, the sampler often needs to first find the most
probable region in the target space; this is a common process often referred to
as ``burn-in'' in the literature. As a consequence, one would discard a certain
amount of data generated from the beginning of the sampling process, and only
use the later part of the chain to estimate the quantities of interest. The
burn-in phase of a gradient-based sampler is often as long as the sampling
phase, which means that a good portion of the computation is not directly
devoted to estimating the target quantities.

% Crux of normalizing flow
All the above issues can be mitigated by normalizing flows.  Normalizing flows
are a technique based on neural networks that aims at learning a mapping from a
simple distribution, such as a Gaussian, to a complex distribution, often given
in the form of samples \cite{2019arXiv190809257K, 2019arXiv191202762P}. Once
the network is trained, one can evaluate the probability density of the complex
distribution and sample from it very efficiently, by first evaluating the
simple distribution and then applying the learned mapping. The core equation
of normalizing flows is the coordinate transformation of probability
distributions via a Jacobian, as given by
\begin{align}
    p_x(X) = p_z(Z) \left| \frac{\partial f}{\partial z}\right|,
\end{align}
\mi{is there an inverse missing on the Jacobian factor?}
where $p_x(X)$ is the complex target distribution, $p_z(Z)$ is the simple
latent distribution and $f$ is an invertible parameterized transform that
connects the two distributions, $x = f(z)$, to be learned by the normalizing
flow. For a detailed discussion of the algorithm, we refer the readers to
\cite{2019arXiv190809257K, 2019arXiv191202762P}.

% The flowMC algorithm
\mi{I think this following paragraph could be made clearer: what does it mean
to learn the ``global landscape''? and how does this jive with the fact that
norm flows are supposed to help with the \emph{local} correlations?}
Working in tandem, gradient-based MCMC and normalizing flows can efficiently
explore posteriors with local and global correlations, as well as multiple
separate modes.  The scheme relies on iteratively using draws from the
gradient-based MCMC to train a normalizing flow, which is then itself used as a
proposal for another stage of MCMC sampling.

Concretely, we begin by producing initial training data for the normalizing
flow by running multiple independent chains of the gradient-based algorithm for
a fixed number of steps.
From the resulting pool of
samples, the normalizing
flow can begin to learn the global landscape of the target distribution.
%
However, since the independent chains contain the same number of samples, the
relative weight assigned to each chains will differ from the true target distribution (e.g., the relative
importance of separate modes will not be correctly calibrated). This is
mitigated by a second stage of gradient-based MCMC sampling that uses the
distribution learned by the normalizing flow as a \textit{proposal}.

Given a trained normalizing flow model, we can generate the proposed jump in
the target space by sampling from the latent distribution $z \sim p_z(Z)$,
usually a Gaussian, and then pushing it through the learned map given by the
normalizing flow model $x=f(z)$.  The acceptance criterion is then set to be
\begin{align}
    \alpha(\mathbf{x},\mathbf{x_0}) = \min \left[ 1, \frac{\hat{\rho}(\mathbf{x_0})\rho_*(\mathbf{x})}{\hat{\rho}(\mathbf{x})\rho_*(\mathbf{x_0})}\right],
\end{align}
where $\hat{\rho}$ is the probability density estimated by the normalizing flow
model, $\rho_*$ is the probability density evaluated using the target function,
and $x_0$ is the current position.

We can see the flow distribution is the
target distribution when the accepting probability is 1. When the normalizing
flow model has not converged to the target distribution, only a portion of the
proposed jumps will be accepted. This means an MCMC process using the
normalizing flow model as the proposal distribution can adjust the normalization
across different regions of the target parameter space by rejecting jumps into
less likely regions. The training and sampling are then repeated until certain
criteria are met.

Note that every time we retrain the network, we are breaking
the Markov properties since we are changing the proposal distribution. To
produce final samples that can be used to estimate target quantities, one has to
freeze the normalizing flow model and not retrain during the final sampling
phase in order to satisfy the detailed balance condition. We use the package
\textsc{flowMC} \cite{2022arXiv221106397W}. The pseudocode of the algorithm is
given in Algorithm \ref{alg:cap}.

\begin{algorithm}
\caption{\textsc{flowMC} pseudocode \mi{can I edit this?}}\label{alg:cap}
\KwIn{initial position $ip$}
\Parameters{n training loop $nt$, n production loop $np$}
\Variables{current chain $cc$, current position $cp$, current NF parameters $\Theta$}
\KwResult{Chains}
$cp \leftarrow ip$\\
\tcc{Training loop}
    \For{$i<nt$}{
        $cc, cp \leftarrow LocalSampling(cp)$\\
        $\Theta \leftarrow TuneNF(cc)$\\
        $chains, cp \leftarrow GlobalSampling(cp, \Theta)$ \\
        $cc \leftarrow Append(cc, chains)$
    }
\tcc{Production loop}
    \For{$i<np$}{
        $c_{local}, cp \leftarrow LocalSampling(cp)$\\
        $c_{global}, cp \leftarrow GlobalSampling(cp, \Theta)$ \\
        $Chains \leftarrow Append(Chains, c_{local}, c_{global})$
    }

@@

Modern \mi{hardware?} accelerators such as graphics processing units (GPUs) and
tensor processing units (TPUs) are designed to execute large-scale dense
computation. They are often much more cost-efficient than using many central
processing units (CPUs) when it comes to solving problems that can be benefited
from parallelization.  The downside of these accelerators compared to CPUs is
that they can only perform a more restricted set of operations and are often
less performant when dealing with serial problems. Parameter estimation with
MCMC is a serial problem since each new sample generated from a chain depends
on the last sample in the chain. This means that naively putting the problem on
an accelerator is more likely to harm performance than improve it.

% More sample helps training normalizing flow
Yet, in our work, the use of accelerators provides two independent perks that
tremendously benefit the parameter estimation process. First, using
accelerators allow us to run many independent MCMC chains simultaneously, which
benefits the training of the normalizing flow. Since we generate the data we
use to train the normalizing flow on the fly, the more independent data we can
feed to the training process, the higher chance the normalizing flow can learn
a reasonable representation of the global landscape of the target distribution.
If we only used a small number of chains, we would be limited to the correlated
samples from each chain and we would have to run more sequential steps to get
the same amount of independent samples---with more chains the problem becomes
parallelizable and we can obtain the same number of training samples sooner. In
other words, being able to use many independent chains helps the normalizing
flow learn the global landscape faster in wall time.

% GPU helps packing a shit ton of waveform evaluation
Another benefit of accelerators is the parallel evaluation of waveforms. Since
the waveform models used in PE are approximants, they can be evaluated at any
given time or frequency independently \mi{is this true for SEOB, and for the
surrogate?}. This means computing a waveform can be trivially parallelized over
frequency bins. Together with the heterodyned likelihood, we can evaluate the
likelihood at $\mathcal{O}(10^7)$ different locations on an Nvidia A100 GPU
\mi{faster?}.  The high throughput of likelihood evaluations unlocks the
potential of the \texttt{flowMC} sampling algorithms.
